---
title: "R Notebook"
author: "Alejandro D. Osborne"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---


```{r}
library(knitr)
library(ggplot2)
library(dplyr)
library(scales)
library(forcats)
```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

We want to see all industries as well as the amount of rows that are encompassed within those industries -

```{r}
knitr::kable(group_by(inc, Industry) %>% summarize(Count=n()) %>% arrange(desc(Count)))
nrow(inc)
```

We're given that this ranks the top 5000 companies, but nrows tells us more is there! Let's hope we don't have to sift through the entire database to find out why, I am going to skip to the end and see what happens - 

```{r}
tail(inc)
```

Perfect (and quite luckily)...there are two companies ranked at 5000 so that kills that mystery. Now I'm most interested in the growth rate perspective, more importantly, just how the distribution falls numerically.

```{r}
rateofGrowth = seq(min(inc$Growth_Rate),max(inc$Growth_Rate),by=(max(inc$Growth_Rate) - min(inc$Growth_Rate))/10)
GrowthRaterange = paste(head(rateofGrowth,-1), rateofGrowth[-1], sep=" - ")
GRFrequency = hist(inc$Growth_Rate, breaks=rateofGrowth, include.lowest=TRUE, plot=FALSE)
data.frame(range = GrowthRaterange, frequency = GRFrequency$counts)
```

This tells us that 98.5% of the growth taken place is under 43%.

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
g <- ggplot(inc, aes(State))
g + geom_bar(aes(fct_infreq(factor(State)), fill=State), position = position_stack(reverse = TRUE), show.legend = F) + coord_flip() + ylab("Companies in Top 5,000") + ggtitle("5,000 Fastest Growing Companies - State")

```


## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
nyGrowth <- subset(inc,State=="NY")
ny <- ggplot(nyGrowth, aes(nyGrowth$Industry, nyGrowth$Employees))
ny + geom_boxplot(na.rm = TRUE) + ylim(0,1000) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
industry <- group_by(inc, Industry)
RevbyEmp <- summarize(industry, revpempl = sum(Revenue)/sum(Employees))
r <- ggplot(RevbyEmp, aes(RevbyEmp$Industry, RevbyEmp$revpempl))
r + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_y_continuous(labels = scales::dollar) + xlab("Industry") + ylab("Revenue per Employee")
```


